---
title: "Lab6_penguins"
author: "Jeremy Knox"
date: "11/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Do mean flipper lengths differe significantly for Adelie, Gentoo and Chinstrap penguins near Palmer Archipeligo?

```{r}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```


Explore the data:

```{r, include=FALSE}
# table for easy counts over multiple factors 
counts = with(penguins, table(Species, Sex))
View(counts)

# update species info with common names using case_when()
penguins2 = penguins %>% 
  mutate(
    common_name = case_when(
      Species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
      Species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
      Species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
    )
  ) %>% 
  select(-Species) %>% 
  filter(Sex == "FEMALE") 
```


Exploratory graphs:
```{r, include=FALSE}
hist_flipper_length = ggplot(penguins2, aes(`Flipper Length (mm)`)) +
  geom_histogram(bins = 10) +
  facet_wrap(~ common_name, scale = "free")
hist_flipper_length
#somewhat bell shaped 

qqs = ggplot(penguins2, aes(sample = `Flipper Length (mm)`)) +
  geom_qq(bins = 10) +
  facet_wrap(~ common_name, scale = "free")
qqs
#looks linear => noramly distributed 

#how do they look compared to eachother?
box = ggplot(penguins2, aes(common_name, `Flipper Length (mm)`)) +
  geom_boxplot(width = .2) +
  geom_jitter(width = 0.1, alpha = 0.5, aes(color = common_name))
box
```


Levene's test for equal variances (keep in mind the general rule: if the greatest variance is less than 4 times bigger than the smallest variance, then usually those are "close enough" to assume equal variance)

```{r, include=FALSE}
variances = penguins2 %>% 
  group_by(common_name) %>% 
  summarize(
    mean = mean(`Flipper Length (mm)`),
    sd = sd(`Flipper Length (mm)`),
    variance = var(`Flipper Length (mm)`)
  )
variances
#smallest = 15, largest = 33, largest < 4*smallest 

penguin_levene = leveneTest(`Flipper Length (mm)` ~ common_name, data = penguins2)
penguin_levene
```


One-Way ANOVA (Type III)
```{r, include=FALSE}
penguin_aov = aov(`Flipper Length (mm)` ~ common_name, data = penguins2)
penguin_sum = summary(penguin_aov)

penguin_ph = TukeyHSD(penguin_aov)

#Testing other opetions (for variation that allows to call values using in-line references...) There's one aov.test() in the new "onewaystests", BUT that doesn't have Tukey's for a podt-hoc pairwise comparison...
penguin_factor = data.frame(
  penguins2 %>% 
  mutate(name_factor = as.factor(common_name),
         flip_numeric = `Flipper Length (mm)`)
)
aov2 = aov.test(flip_numeric ~ name_factor, data = penguin_factor)
#Yes, there are at least two means that differ signiificantly.
#Which ones?

#Bonferrono correction:
post_hoc = paircomp(aov2, adjust.method = "bonferroni")
#All means are significantly different from each other!
```

How do we report this outcome?

Mean flipper lengths for Adelie (mean ± SD, n = ), Gentoo (mean ± sd, n = ), and Chinstrap (mean ± sd, n =) penguins all differed significantly across all groups by one-way ANOVA (F(`r aov2$parameter`) = `r aov2$statistic`, $p$ < 0.001, $\alpha$ = 0.05) with post-hox pairwise comparison by Bonferroni correction ($p$ < 0.001 for all pairwise comparisons)









